Indicators for Ecosystem Condition in Norway

### 2023-10-12 functional_plant_indicators_wetland.Rmd

@@

<!-- Load all you dependencies here -->
```{r setup, include=FALSE}
library(sf)
library(plyr)
library(tidyverse)
library(RColorBrewer)
library("gridExtra") 
library(ggridges)
library(tmap)
library(tmaptools)
library(raster)
library(stars)
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

@@

```{r, echo = F}
# Add data from cache
ANO.sp<-readRDS(paste0(here::here(),"/data/cache/ANO.sp.RDS"))
ANO.geo<-readRDS(paste0(here::here(),"/data/cache/ANO.geo.RDS"))

@@

Plant indicators from Tyler et al. (2021)
```{r, echo = F}
## Tyler indicator data
## data from cache
ind.dat <-readRDS(paste0(here::here(),"/data/cache/ind.dat.RDS"))
# ind.dat <- read.table("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/functional plant indicators/Tyler et al_Swedish plant indicators.txt",
#                         sep = '\t', header=T, quote = '')
# head(ind.dat)

@@

Generalized species lists NiN
```{r, echo = F}
# data from cache
Eco_State<-readRDS(paste0(here::here(),"/data/cache/Eco_State.RDS"))
# load("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/functional plant indicators/reference from NiN/Eco_State.RData")
# str(Eco_State)
```

@@

```{r, include = T, attr.output='style="max-height: 300px;"'}

@@

NiN.wetland.cov$sp <- as.factor(NiN.wetland.cov$sp)
```
leaving us with the monitoring data including plant indicators (ANO.sp.ind) and the reference data including plant indicators (NiN.wetland.cov)

<!-- Print head of data set with horizontal scrolling -->
```{r}
head(ANO.sp.ind) %>%
  kable("html") %>% kable_styling("striped") %>% scroll_box(width = "100%")
```

```{r}
head(NiN.wetland.cov) %>% 
  kable("html") %>% kable_styling("striped") %>% scroll_box(width = "100%")
```

@@

##### bootstrap function for frequency abundance
- function to calculate community weighted means of selected indicator values (ind)
- for species lists (sp) with given abundances in percent (or on a scale from 0 to 1) in one or more 'sites' (abun)
- with a given number of iterations (iter),
- with species given a certain minimum abundance occurring in all bootstraps (obl), and
- with a given re-sampling ratio of the original species list (rat)
- in every bootstrap iteration the abundance of the sampled species can be randomly changed by a limited amount if wished by introducing a re-sampling of abundance values from adjacent abundance steps with a certain probability (var.abun)

@@

Running the bootstraps

@@

```{r, eval = F}
wetland.ref.cov <- indBoot.freq(sp=NiN.wetland.cov[,1],abun=NiN.wetland.cov[,2:46],ind=NiN.wetland.cov[,47:50],
                          iter=1000,obl=1,rat=2/3,var.abun=T)

@@

```{r}
head(wetland.ref.cov[[1]]) %>%
    kable("html") %>% kable_styling("striped") %>% scroll_box(width = "100%")

@@

This results in an R-list with a slot for every selected indicator, and in every slot there's a data frame with as many columns as there are NiN species lists and as many rows as there were iterations in the bootstrap.
Next, we need to derive scaling values from these bootstrap-lists (the columns) for every mapping unit in NiN. Here, we define things in the following way:

@@

```{r, attr.output='style="max-height: 300px;"'}

@@

```{r}
head(wetland.ref.cov.val) %>%
      kable("html") %>% kable_styling("striped") %>% scroll_box(width = "100%")

@@

Here is the scaling function
```{r}

@@

We then can prepare a list of data frames to hold the results and perform the scaling according to the principles described in NINA report 1967 (Töpper and Jakobsson 2021)
```{r}

@@

```{r, echo = T, eval = F}

@@

```{r}
head(results.wetland[['2-sided']]) %>%
      kable("html") %>% kable_styling("striped") %>% scroll_box(width = "100%")
```
#### Scaled value analyses

@@

### Ecosystem sub-types {#ecosystem-sub-types-fpiw}
And we can show the resulting scaled values as Violin plots for each indicator and main ecosystem type 
```{r fpiw-plot-maintype,echo=FALSE,message=FALSE, warning=FALSE, fig.cap="Scaled vegetation indicator values for wetlands. 1 = cases where variable is lower than the reference value- 2 = cases where variablei si higher than the reference value."}

# making the plot
ggplot(res.wetland, aes(x=factor(hovedtype_rute), y=scaled_value, fill=fp_ind)) + 
  geom_hline(yintercept=0.6, linetype="dashed") + 
  geom_violin() +
#  geom_boxplot(width=0.2, color="grey") +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap(~factor(fp_ind,levels=c("Light1","Moist1","pH1","Nitrogen1","Light2","Moist2","pH2","Nitrogen2")), ncol = 4) + 
  xlab("Main ecosystem type") + 
  ylab("Scaled indicator value") 
```

@@

```{r ph-lower-map-fpiw, echo=FALSE, warning=FALSE,message = FALSE, fig.cap="Functional plant indicator values for pH (lower) in wetlands."}
#### scaled value maps ####
# keep wide format and add geometry again
res.wetland2 <- results.wetland[['2-sided']]
st_geometry(res.wetland2) <- st_geometry(ANO.wetland)

@@

Therefore, we need to (i) use a beta-model, that (ii) can account for the nested structure of the data.
Here, we apply the following function using either a glmmTMB null-model with a beta-distribution, logit link, and the nesting as a random intercept, or a simple betareg null-model with logit link if the nesting is not extensive enough for a mixed model.
```{r}
library(betareg)
library(glmmTMB)

expit <- function(L) exp(L) / (1+exp(L)) # since the beta-models use a logit link, we need to calculate the estimates back to the identity scale

@@

```{r}

@@

## scaled value maps for pH1 (lower indicator)
# mean
tm_shape(regnor) +
  tm_polygons(col="pH1.reg.mean", title="pH (lower), mean", style="quantile", palette=rev(get_brewer_pal(palette="OrRd", n=5, plot=FALSE))) +
  tm_text("pH1.reg.n",col="black",bg.color="grey")
```
Mean index value by region for the lower pH indicator (i.e. index shows deviations towards a more acidic environment). Numbers in grey fields show the number of observations in the respective region.

```{r}
# se
tm_shape(regnor) +
  tm_polygons(col="pH1.reg.se", title="pH (lower), sd", style="quantile", palette=get_brewer_pal(palette="OrRd", n=5, plot=FALSE)) +
  tm_text("pH1.reg.n",col="black",bg.color="grey")

```
Standard error to the mean index value by region for the lower pH indicator. Numbers in grey fields show the number of observations in the respective region.

### unscaled values vs. reference
We can also compare the unscaled values to the reference distribution in order to identify ecosystem types and functional plant indicators showing a deviation from the expectation. Since pH and nitrogen show some deviation in the lower indicator we exemplify this with these two indicators for unscaled values.

```{r}
summary(res.wetland$kartleggingsenhet_1m2)
length(unique(res.wetland$kartleggingsenhet_1m2))
# 16 NiN-types to plot
colnames(wetland.ref.cov[['Soil_reaction_pH']])

@@

# V1-C-8
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V1-C8a","V1-C8b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C8',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original)),
       col="red")

par(mfrow=c(2,4))
## V2s
# V2-C-1
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V2-C1a","V2-C1b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V2-C1',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original)),
       col="red")

@@

# V8-C-2
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V8-C2")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V8-C2',xlab='pH value')
#points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V8-C-2",]$original,na.rm=T),
#       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V8-C-2",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V8-C-2",]$original)),
       col="red")
legend("topleft", legend=c("reference","field data"), pch=c(NA,1), lty=1, col=c("black","red"))

### Nitrogen
par(mfrow=c(2,4))
## V1s
# V1-C-1
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V1-C1a","V1-C1b","V1-C1c","V1-C1d","V1-C1e")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C1',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-1",]$original)),
       col="red")

@@

# V1-C-8
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V1-C8a","V1-C8b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C8',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original)),
       col="red")

par(mfrow=c(2,4))
## V2s
# V2-C-1
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V2-C1a","V2-C1b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V2-C1',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original)),
       col="red")

@@

legend("topleft", legend=c("reference","field data"), pch=c(NA,1), lty=1, col=c("black","red"))

```
Deviations from the reference occur mainly in V2-C1, which represents limestone-poor swamp forests. According to the functional signature from the plant community composition many of the occurrences of this ecosystem type are too acidic and may have too low availability of nitrogen.

### Eksport file (final product) {#exp-fpiw}
<!-- Export final file. Ideally a georeferenced shape or raster wit indicators values (raw and normalised), reference values and errors. -->

>>>>

---
output:
  html_document:
    code_folding: show
---

@@

<!-- Load all you dependencies here -->
```{r setup, class.source = 'fold-hide'}
library(sf)
library(plyr)
library(tidyverse)
library(RColorBrewer)
library(gridExtra) 
library(ggridges)
library(tmap)
library(tmaptools)
library(raster)
library(stars)
library(knitr)
library(kableExtra)
library(betareg)
library(glmmTMB)
knitr::opts_chunk$set(echo = TRUE)
```

<!-- Limit code block height and force scrolling -->
```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
}
```

@@

```{r, class.source = 'fold-hide'}
# Add data from cache
ANO.sp<-readRDS(paste0(here::here(),"/data/cache/ANO.sp.RDS"))
ANO.geo<-readRDS(paste0(here::here(),"/data/cache/ANO.geo.RDS"))

@@

Plant indicators from Tyler et al. (2021)
```{r, class.source = 'fold-hide'}
## Tyler indicator data
## data from cache
ind.dat <-readRDS(paste0(here::here(),"/data/cache/ind.dat.RDS"))
# ind.dat <- read.table("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/functional plant indicators/Tyler et al_Swedish plant indicators.txt",
#                         sep = '\t', header=T, quote = '')
# head(ind.dat)

@@

Generalized species lists NiN
```{r, class.source = 'fold-hide'}
# data from cache
Eco_State<-readRDS(paste0(here::here(),"/data/cache/Eco_State.RDS"))
# load("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/functional plant indicators/reference from NiN/Eco_State.RData")
# str(Eco_State)
```

@@

```{r, include = T, results='hide', warning=F, message=F, class.source = 'fold-hide'}

@@

NiN.wetland.cov$sp <- as.factor(NiN.wetland.cov$sp)
```

This leaves us with the monitoring data including plant indicators (ANO.sp.ind) and the reference data including plant indicators (NiN.wetland.cov):

<!-- Print head of data set with horizontal scrolling -->
```{r ano-sp-tab, class.source = 'fold-hide'}
head(ANO.sp.ind) %>%
  kable("html",caption = "ANO species occurence data set with attached plant trait data.") %>% 
  kable_styling("striped") %>% scroll_box(width = "100%")
```

```{r wetland-ref-data-tab, class.source = 'fold-hide'}
head(NiN.wetland.cov) %>% 
  kable("html", caption = "Reference data set") %>% 
  kable_styling("striped") %>% scroll_box(width = "100%")
```

@@

##### Bootstrap function for frequency abundance
- function to calculate community weighted means of selected indicator values (ind)
- for species lists (sp) with given abundances in percent (or on a scale from 0 to 1) in one or more 'sites' (abun)
- with a given number of iterations (iter),
- with species given a certain minimum abundance occurring in all bootstraps (obl), and
- with a given re-sampling ratio of the original species list (rat)
- in every bootstrap iteration the abundance of the sampled species can be randomly changed by a limited amount if wished by introducing a re-sampling of abundance values from adjacent abundance steps with a certain probability (var.abun)

@@

@@

Running the bootstraps:
```{r, eval = F}
wetland.ref.cov <- indBoot.freq(sp=NiN.wetland.cov[,1],abun=NiN.wetland.cov[,2:46],ind=NiN.wetland.cov[,47:50],
                          iter=1000,obl=1,rat=2/3,var.abun=T)

@@

```{r}
head(wetland.ref.cov[[1]]) %>%
    kable("html", caption = "Table showing the first 6 rows of the boostrapped data set.") %>% kable_styling("striped") %>% scroll_box(width = "100%")

@@

This results in an R-list with a slot for every selected indicator, and in every slot there's a data frame with as many columns as there are NiN species lists and as many rows as there were iterations in the bootstrap.

Next, we need to derive scaling values from these bootstrap-lists (the columns) for every mapping unit in NiN. Here, we define things in the following way:

@@

```{r, attr.output='style="max-height: 300px;"', results='hide', class.source = 'fold-hide'}

@@

```{r}
wetland.ref.cov.val %>%
      kable("html", caption = "Reference values (Rv and threshold values (Gv) for each indicator and nature type combination") %>% kable_styling("striped") %>% scroll_box(width = "100%", height = "300px")

@@

Here is the scaling function:
```{r}

@@

We then can prepare a list of data frames to hold the results and perform the scaling according to the principles described in NINA report 1967 (Töpper and Jakobsson 2021).
```{r, results='hide', class.source = 'fold-hide'}

@@

```{r, eval = F, class.source = 'fold-hide'}

@@

```{r}
head(results.wetland[['2-sided']]) %>%
      kable("html", caption = "ANO plots for wetland ecosystems with scaled indicator values in the far-right columns.") %>% 
  kable_styling("striped") %>% scroll_box(width = "100%")
```
#### Scaled value analyses

@@

### Ecosystem sub-types {#ecosystem-sub-types-fpiw}
And we can show the resulting scaled values as Violin plots for each indicator and main ecosystem type 
```{r fpiw-plot-maintype, message=FALSE, warning=FALSE, fig.cap="Scaled vegetation indicator values for wetlands. 1 = cases where variable is lower than the reference value- 2 = cases where variablei is higher than the reference value.", out.width='100%', class.source = 'fold-hide'}

# making the plot
ggplot(res.wetland, aes(x=factor(hovedtype_rute), y=scaled_value, fill=fp_ind)) + 
  geom_hline(yintercept=0.6, linetype="dashed") + 
  geom_violin() +
#  geom_boxplot(width=0.2, color="grey") +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap(~factor(fp_ind,levels=c("Light1","Moist1","pH1","Nitrogen1","Light2","Moist2","pH2","Nitrogen2")), ncol = 4) + 
  xlab("Wetland ecosystem sub-type") + 
  ylab("Scaled indicator value") 
```

This figure is a bit easier to read if we remove the sub-types with very few data points:
```{r fpiw-plot-maintype-v2, message=FALSE, warning=FALSE, fig.cap="Scaled vegetation indicator values for the four most common sub-types of wetland. 1 = cases where variable is lower than the reference value- 2 = cases where variablei is higher than the reference value.", out.width='100%', class.source = 'fold-hide'}
ggplot(res.wetland[res.wetland$hovedtype_rute %in% c("V1", "V2", "V3", "V4"),], aes(x=factor(hovedtype_rute), y=scaled_value, fill=fp_ind)) + 
  geom_hline(yintercept=0.6, linetype="dashed") + 
  geom_violin() +
#  geom_boxplot(width=0.2, color="grey") +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap(~factor(fp_ind,levels=c("Light1","Moist1","pH1","Nitrogen1","Light2","Moist2","pH2","Nitrogen2")), ncol = 4) + 
  xlab("Wetland ecosystem sub-type") + 
  ylab("Scaled indicator value") +
  guides(fill='none')
```

@@

```{r ph-lower-map-fpiw, warning=FALSE, message = FALSE, fig.cap="Functional plant indicator values for pH (lower) in wetlands.", class.source = 'fold-hide'}
#### scaled value maps ####
# keep wide format and add geometry again
res.wetland2 <- results.wetland[['2-sided']]
st_geometry(res.wetland2) <- st_geometry(ANO.wetland)

@@

Therefore, we need to (i) use a beta-model, that (ii) can account for the nested structure of the data.
Here, we apply the following function using either a glmmTMB null-model with a beta-distribution, logit link, and the nesting as a random intercept, or a simple betareg null-model with logit link if the nesting is not extensive enough for a mixed model.
```{r}
expit <- function(L) exp(L) / (1+exp(L)) # since the beta-models use a logit link, we need to calculate the estimates back to the identity scale

@@

```{r ph-map-fpiw, fig.cap="Mean index value by region for the lower pH indicator (i.e. index shows deviations towards a more acidic environment). Numbers in grey fields show the number of observations in the respective region."}

@@

# Colour palettes
myCol <- "RdYlGn"

## scaled value maps for pH1 (lower indicator)
# mean
tm_shape(regnor) +
  tm_polygons(col="pH1.reg.mean", title="pH (lower), mean", style="fixed", 
              palette=myCol,
              breaks = seq(0,1,.2)) +
  tm_text("pH1.reg.n",col="black",bg.color="grey")
```

We can also plot the uncertainty, here represented as the spatial variation:
```{r ph-se-fpiw, fig.cap="Standard error to the mean index value by region for the lower pH indicator. Numbers in grey fields show the number of observations in the respective region."}

tm_shape(regnor) +
  tm_polygons(col="pH1.reg.se", title="pH (lower), sd", style="quantile", palette=get_brewer_pal(palette="OrRd", n=5, plot=FALSE)) +
  tm_text("pH1.reg.n",col="black",bg.color="grey")

### Unscaled values vs. reference
We can also compare the unscaled values to the reference distribution in order to identify ecosystem types and functional plant indicators showing a deviation from the expectation. Since pH and nitrogen show some deviation in the lower indicator we exemplify this with these two indicators for unscaled values.

```{r pH-distcomp-fpiw, class.source = 'fold-hide', fig.cap="Comparing the reference and variable distribution for the pH indicator in ecosystem sub-type V1."}
    #summary(res.wetland$kartleggingsenhet_1m2)
    #length(unique(res.wetland$kartleggingsenhet_1m2))
# 16 NiN-types to plot
    #colnames(wetland.ref.cov[['Soil_reaction_pH']])

@@

# V1-C-8
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V1-C8a","V1-C8b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C8',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original)),
       col="red")
```

```{r ph-distcomp2-fpiw, class.source = 'fold-hide', fig.cap="Comparing the reference and variable distribution for the pH indicator in ecosystem sub-types other than V1."}
par(mfrow=c(2,4))
## V2s
# V2-C-1
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V2-C1a","V2-C1b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V2-C1',xlab='pH value')
points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original)),
       col="red")

@@

# V8-C-2
plot(density( as.matrix(wetland.ref.cov[['Soil_reaction_pH']][,c("V8-C2")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V8-C2',xlab='pH value')
#points(density(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V8-C-2",]$original,na.rm=T),
#       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V8-C-2",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="pH1" & res.wetland$fp_ind=="pH1" & res.wetland$kartleggingsenhet_1m2=="V8-C-2",]$original)),
       col="red")
legend("topleft", legend=c("reference","field data"), pch=c(NA,1), lty=1, col=c("black","red"))
```

```{r n-distcomp-fpiw, class.source = 'fold-hide', fig.cap="Comparing the reference and variable distribution for the nitrogen affiliation indicator in ecosystem sub-type V1."}
### Nitrogen
par(mfrow=c(2,4))
## V1s
# V1-C-1
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V1-C1a","V1-C1b","V1-C1c","V1-C1d","V1-C1e")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C1',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-1",]$original)),
       col="red")

@@

# V1-C-8
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V1-C8a","V1-C8b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V1-C8',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V1-C-8",]$original)),
       col="red")
```

```{r n-distcomp2-fpiw,class.source = 'fold-hide', fig.cap="Comparing the reference and variable distribution for the nitrogen affiliation indicator in ecosystem sub-types other than V1."}
par(mfrow=c(2,4))
## V2s
# V2-C-1
plot(density( as.matrix(wetland.ref.cov[['Nitrogen']][,c("V2-C1a","V2-C1b")]) ,na.rm=T),
     xlim=c(1,7), type="l", main='V2-C1',xlab='Nitrogen value')
points(density(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original,na.rm=T),
       type="l", col="red")
points(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original,
       rep(0,length(res.wetland[res.wetland$fp_ind=="Nitrogen1" & res.wetland$fp_ind=="Nitrogen1" & res.wetland$kartleggingsenhet_1m2=="V2-C-1",]$original)),
       col="red")

@@

legend("topleft", legend=c("reference","field data"), pch=c(NA,1), lty=1, col=c("black","red"))
```

Deviations from the reference occur mainly in V2-C1, which represents limestone-poor swamp forests. According to the functional signature from the plant community composition many of the occurrences of this ecosystem type are too acidic and may have too low availability of nitrogen.

### Eksport file (final product) {#exp-fpiw}
Nothing exported yet.
